---
title: "3.2 Querying CYorks OMOP"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

library(ggplot2)
library(dplyr)
library(DBI)
library(knitr)

get_omop_data <- function(sql) {
  
  con <- dbConnect(odbc::odbc(), .connection_string = "Driver={SQL Server Native Client 10.0};Server=BHTS-CONYPROD16;Trusted_Connection=yes;Database=CDM_TEST")

  dat <- dbGetQuery(con, sql)

  dbDisconnect(con)
  
  return(dat)
  
}

```
```{r getdata}

sql <- "select person_id, year_of_birth from person;"
person_year <- get_omop_data(sql)

sql <- "select person_id, gender_concept_id from person;"
person_gender <- get_omop_data(sql)

sql <- "select concept_id, concept_name from concept where concept_id in (8507, 8532, 8551, 45454912);"
gender_concept <- get_omop_data(sql)

sql <- "select person_id, concept_name
from person
inner join concept
on gender_concept_id = concept_id;"
person_gender_labelled <- get_omop_data(sql)

```

## Person

Basic demographic information is held in the `person` table.

```{sql eval = FALSE, echo = TRUE}
select person_id, year_of_birth
from person;
```

```{r echo = TRUE}

ggplot(person_year, aes(x = year_of_birth)) + geom_bar()

```

Everything in OMOP is a concept stored in the concept table. When we query many of the data items directly, we just see the concept ids.

```{sql eval = FALSE, echo = TRUE}
select person_id, gender_concept_id
from person;
```

```{r echo=TRUE}

person_gender <- person_gender %>% 
  group_by(gender_concept_id) %>% 
  summarise(n = n())

kable(person_gender, align = "l")

```

If we want to find out what the concept ids mean, we need to look in the concept table. The definition of the concept is found in the `concept_name` field.

```{sql eval = FALSE, echo = TRUE}
select concept_id, concept_name 
from concept 
where concept in (8507, 8532, 8551, 45454912);
```

```{r echo=TRUE}

kable(gender_concept, align = "l")

```

We can also join the `concept_name` field directly onto our gender query.

```{sql eval = FALSE, echo = TRUE}
select person_id, concept_name
from person
inner join concept
on gender_concept_id = concept_id;
```

```{r echo=TRUE}

person_gender_labelled <- person_gender_labelled %>% 
  group_by(concept_name) %>% 
  summarise(n = n())

kable(person_gender_labelled, align = "l", col.names = c("Gender", "count"))

```
