---
title: "3.2 Querying CYorks OMOP"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

library(ggplot2)
library(dplyr)
library(DBI)
library(knitr)

get_omop_data <- function(sql) {
  
  con <- dbConnect(odbc::odbc(), .connection_string = "Driver={SQL Server Native Client 10.0};Server=BHTS-CONYPROD16;Trusted_Connection=yes;Database=CDM_TEST")

  dat <- dbGetQuery(con, sql)

  dbDisconnect(con)
  
  return(dat)
  
}

```
```{r getdata}

sql <- "select person_id, year_of_birth from person;"
person_year <- get_omop_data(sql)

sql <- "select person_id, gender_concept_id from person;"
person_gender <- get_omop_data(sql)

sql <- "select concept_id, concept_name from concept where concept_id in (8507, 8532, 8551, 45454912);"
gender_concept <- get_omop_data(sql)

sql <- "select person_id, concept_name
from person
inner join concept
on gender_concept_id = concept_id;"
person_gender_labelled <- get_omop_data(sql)

sql <- "select death_date, count(*) as n from death group by death_date order by death_date;"
death <- get_omop_data(sql)

sql <- "select c.concept_name as encounter_type, count(*) as n from visit_occurrence v 
inner join concept c on v.visit_concept_id = c.concept_id
group by c.concept_name;"
visit <- get_omop_data(sql)


```

## Person

Basic demographic information is held in the `person` table.

```{sql eval = FALSE, echo = TRUE}
select person_id, year_of_birth
from person;
```

```{r echo = TRUE}

ggplot(person_year, aes(x = year_of_birth)) + geom_histogram(binwidth = 1) + theme_classic()

```

Everything in OMOP is a concept stored in the concept table. When we query many of the data items directly, we just see the concept ids.

```{sql eval = FALSE, echo = TRUE}
select person_id, gender_concept_id
from person;
```

```{r echo=TRUE}

person_gender <- person_gender %>% 
  group_by(gender_concept_id) %>% 
  summarise(n = n())

kable(person_gender, align = "l")

```

If we want to find out what the concept ids mean, we need to look in the concept table. The definition of the concept is found in the `concept_name` field.

```{sql eval = FALSE, echo = TRUE}
select concept_id, concept_name 
from concept 
where concept in (8507, 8532, 8551, 45454912);
```

```{r echo=TRUE}

kable(gender_concept, align = "l")

```

We can also join the `concept_name` field directly onto our gender query.

```{sql eval = FALSE, echo = TRUE}
select person_id, concept_name
from person
inner join concept
on gender_concept_id = concept_id;
```

```{r echo=TRUE}

person_gender_labelled <- person_gender_labelled %>% 
  group_by(concept_name) %>% 
  summarise(n = n())

kable(person_gender_labelled, align = "l", col.names = c("Gender", "count"))

```

Deaths information is held in a separate table. The following `sql` code computes the number of deaths on each day. `order by` is needed for the cumulative sum function `cumsum(n)` to work as expected.

```{sql eval = FALSE, echo = TRUE}
select death_date, count(*) as n
from death
group by death_date
order by death_date;
```


```{r echo = TRUE}

ggplot(death, aes(x = death_date)) + geom_line(aes(y = cumsum(n))) + theme_classic()

```


## Visits

Visits are documented in the `visit_occurrence` table. Each visit record represents an encounter that may lead to clinical data being recorded about the person. There are four key linkages to be aware of here:

* `person_id` refers to the person's record in the `person` table
* `visit_concept_id` refers to the `concept_id` in the `concept` table describing the type of encounter
* `care_site_id` refers to the `care_site` table where the care site of the encounter is identified
* `visit_occurrence_id` is the primary key for the `visit_occurrence` record and appears in all clinical data records that link to the visit

```{r}

blogdown::shortcode(.name = 'panel', status = 'notice', title = 'Note', 
  .content = 'Visit information can also be captured in the `visit_detail` table. This is not currently used by Connected Yorkshire, although it may be in a future release.')

```

We can use the `visit_occurrence` table and its various linkages to explore different aspects of service use across our dataset. For example, we can count number of visits by type of encounter. To do this, we need to join the `visit_concept_id` to the `concept` table in order to get the relevant `concept_name`.

```{sql eval = FALSE, echo = TRUE}
select c.concept_name as encounter_type, count(*) n from visit_occurrence v 
inner join concept c on v.visit_concept_id = c.concept_id
group by c.concept_name;
```


```{r echo = TRUE}

ggplot(visit, aes(x = encounter_type, y = n)) + geom_col() + theme_classic()

```



